Signatures of Foreground Sky in Power Spectrum of Redshifted Neutral Hydrogen from the Epoch of Reionization

### 2014-09-21 foreground_delay_spectrum_modeling.tex
From 2dc2fc1: ============
(*) Minor edits from section 4 to the end
(*) Need to verify color scale on delay spectra

We describe power--pattern and foreground models we have used in modeling the instrument. %todo

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
\caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the off--zenith (top) and zenith pointings (bottom) at 185~MHz. An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above the ground plane. Random fluctuations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators. The logarithmic color scale (shown at the right) is common to both panels. \label{fig:power_pattern}}
\end{figure}

@@

The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non--zero response all the way to the horizon. As we have seen above, foreground objects contaminate the region inside the {\it wedge} with an amplitude that depends on the amplitude of power--pattern in that direction, with objects at the horizon being closest to the boundaries of the wedge.

The degree to which these objects leak power into spectral line of sight modes above the {\it foreground wedge}, depends primarily on the brightness of foregrounds near the boundaries of the {\it wedge} \citep{thy13,pob13,ved12,par12}. Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view.

@@

For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz. Since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz diffuse emission map to the same resolution. However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals. We model the spectral dependence of the diffuse foreground as a single spectral index at each pixel in our map, found by comparing model maps at 170~MHz and 200~MHz.

A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}. Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power receptivity is significantly high ($\gtrsim 12$\%) in that direction. In the {\it zenith} pointing, the Galactic plane has set and the contour level of power--pattern in that direction is at least 16 times smaller.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
\caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels. Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5. The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power receptivity is significant ($\gtrsim 12$\%) in that direction. In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser. \label{fig:DSM}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}. The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. \label{fig:sim_data}}
\end{figure}

In order to make a quantitative comparison of delay spectra obtained with the MWA data and our simulations, we consider the uncertainty in the assumed spectral index of our foreground model. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency. Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact objects from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$70\% and $\sim$30\% respectively.

@@

Figure~\ref{fig:data-sim-ratio} shows histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with aforementioned uncertainty in foreground models and thermal noise fluctuations in measurements. Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Histogram of logarithm of ratio of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases. A value of zero indicates when the ratio $\rho=1$. \label{fig:data-sim-ratio}}
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Probability densities and cumulative distributions are shown in black and gray respectively. Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty. Also indicated in percentages are confidence levels estimated conservatively for such a consistency. They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. \label{fig:data-sim-ratio-snr}}
\end{figure}

These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved. However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power--pattern, calibration, radio frequency interference (RFI), and wide field imaging effects, our estimates are conservative. Accounting for such uncertainties will only improve our confidence level in this comparison. Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

Having established that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in the simulated delay spectra. A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through simulations. We address these factors below:

@@

\item {\it Sky Model}: Our model of the sky consists of bright compact objects and diffuse emission on diverse spatial scales as seen from Figures~\ref{fig:DSM} and \ref{fig:CSM}. It also consists of localized regions of strong emission such as the Galactic plane and Galactic center. The resulting sky emission is anisotropic. In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission. % In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions.  Each region has equal area and the regions are aligned to cardinal directions around zenith as the origin. This results in the ability to isolate features in the delay spectrum to eight different sectors of the sky.

\item {\it Antenna Pair Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna pairs by their orientation, $\theta_\textrm{b}$. We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$. The bin centers are oriented towards South--East, East, North--East, and North respectively. Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match delay spectrum features to different sky directions.

\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the object. The pointing system is capable of steering to points on a regular $\sim$7\arcdeg grid. During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary. In our study, we take into account the effect of changing power--pattern of the antennas on the delay spectrum (Figure~\ref{fig:power_pattern}). The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.

\item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to the Fourier relationship between the two. Due to excision of aliased edge channels repeated in each coarse channel of our passband, we observed repeated patterns of the {\it foreground wedge} in the delay spectra. As already mentioned earlier, we have used the deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects. Since this paper focuses on studying the effects of foreground on the delay spectra, we do not pursue the characterization of artifacts from the band shape and imperfections in {\it CLEAN} deconvolution.

@@

It may be noted that numerous features overlap at varying levels of significance depending on combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna pair orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectrum of foreground emission.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. In the {\it zenith} pointing, delay spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings. \label{fig:noiseless-dsm-delay-spectrum}}
\end{figure}

@@

The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the far west on one of the sidelobes of the power--pattern. It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increase in antenna spacing. This Galactic center signature is almost invisible in the {\it zenith} pointing because the Galactic center has already set over the horizon (see bottom panel in Figure~\ref{fig:noiseless-dsm-delay-spectrum}).

@@

As we also see in Figure~\ref{fig:noiseless-dsm-delay-spectrum}, diffuse emission from within the primary field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, while in the {\it zenith} pointing it shows up at near zero delay. The former is seen at positive delays because the primary lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith. As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to length of the baseline projected in the direction of emission. Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\overline{\mathbf{x}}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75. In the off-zenith pointing, we see power out to much longer baselines which is described below.

@@

A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m)---contrary to expectations, since longer baselines are not expected to be sensitive to diffuse emission.  In the Widefield observations of the MWA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.  This effect is evident at all LSTs in our simulations and it naturally holds for short baselines, as well. Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines. It decreases in strength with increasing baseline length as expected but is present in all orientations of the baseline vector.   %% This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:composite} in more detail.

@@

Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of compact foreground sky model for {\it off--zenith} and {\it zenith} pointings (shown in Figure~\ref{fig:CSM}) in the top and bottom panels respectively.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Simulated delay spectrum amplitude, $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz), obtained with simulations for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and logarithmic color scale are identical to those in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}. Emission from compact objects gives rise to centrally concentrated features in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
\end{figure}

@@

The amplitude response of an interferometer to a compact object is, to first order, flat with respect to baseline length. Therefore, the distribution of objects within the primary field of view depends mainly on how the power--pattern maps to delay space for each antenna pair. Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the primary lobe of the power--pattern. In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the primary lobe of the power--pattern is centered at zenith.

% \subsubsection{Compact Emission in Power--Pattern Sidelobes}\label{sec:csm-side-lobe}

Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with sidelobes of the power--pattern. On the other hand, compact objects co--located with secondary lobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.

@@

% \subsubsection{{\it Pitchfork} Effect}\label{sec:pitchfork}

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional power spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}. The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively. The horizon delay limits are shown as a pair of vertical dot--dashed lines. The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum. Objects inside the {\it foreground} wedge (between the horizon delay limits) are dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, outside the wedge thermal noise dominates.

>>>>

We describe power--pattern and foreground models we have used in our simulations. %todo

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_powerpattern_0.3m_ground_custom.eps}
\caption{Models of MWA tile power--pattern in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates for the {\it off--zenith} (top) and {\it zenith} pointings (bottom) at 185~MHz. An MWA tile is modeled as a 4$\times$4 array of isotropic radiators placed 0.3~m above an infinite ground plane. Random fluctuations of rms 0.05~ns have been added to delay corrections during the phased addition of voltages from these isotropic radiators. The logarithmic color scale (shown on the right) is common to both panels. \label{fig:power_pattern}}
\end{figure}

@@

The MWA has a very wide field of view ($\gtrsim 20$\arcdeg at 185~MHz) and has non--zero response all the way to the horizon. As already noted, foreground objects fill the {\it foreground wedge} with an amplitude which depends on the tile power gain in that direction, with objects at the horizon being closest to the boundaries of the {\it wedge}.

The degree to which these objects leak power into spectral line of sight modes in the {\it EoR window} will be shown to depend primarily on the brightness of foregrounds near the boundaries of the {\it foreground wedge}. Thus, it is important to consider an all--sky model for foreground objects in evaluating the features seen in the power spectrum instead of restricting only to the primary field of view.

@@

For the diffuse component, we use an all--sky radio foreground model \citep{deo08} to estimate the emission at 185~MHz. Since this map is predominantly based on the 408~MHz all--sky map of \citet{has82} which has an angular resolution of 0\fdg 85, we smoothed the 185~MHz map to the same resolution. However, to avoid any artifacts from sampling this map, we sample it at $\approx 27$\arcmin~intervals. We model the diffuse foreground spectra with a single spectral index at each pixel in the map, found by comparing model maps at 170~MHz and 200~MHz.

A low resolution version of the diffuse foreground model used is shown in Figure~\ref{fig:DSM}. Contours of the MWA tile power--pattern shown in Figure~\ref{fig:power_pattern} are also overlaid. Of notable significance in the {\it off--zenith} pointing is the presence of a portion of the Galactic plane and the bright Galactic center in the visible hemisphere in the westward sky and the MWA tile's power gain is significantly high ($\gtrsim 12$\%) in that direction. In the {\it zenith} pointing, the Galactic plane has set and the contour level of power--pattern in that direction is at least 16 times smaller.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/dsm.eps}
\caption{Sky brightness temperature (in K) of the diffuse foreground model at 185~MHz in Right Ascension ($\alpha$) and Declination ($\delta$) coordinates visible during {\it off--zenith} (top) and {\it zenith} (bottom) pointings. The color scale on the right is logarithmic and is common to both panels. Contours of power--pattern shown in Figure~\ref{fig:power_pattern} are overlaid. The contour levels shown are 0.00195, 0.00781, 0.0312, 0.125, and 0.5. The Galactic center and a portion of the Galactic plane are prominently visible during the {\it off--zenith} pointing and the MWA tile power gain is significant ($\gtrsim 12$\%) in that direction. In contrast, emission from the Galactic plane in {\it zenith} pointing is significantly lesser. \label{fig:DSM}}
\end{figure}

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/multi_baseline_sim_delay_spectrum_snapshots.eps}
\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings. White lines mark the boundaries of {\it foreground wedge} determined by horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figure~\ref{fig:fhd_data}. The features resemble those obtained with MWA data shown in Figure~\ref{fig:fhd_data}. The Galactic center and Galactic plane prominently visible in the {\it off--zenith} pointing makes the emission in the {\it foreground wedge} brighter relative to that in the {\it zenith} pointing. \label{fig:sim_data}}
\end{figure}

In order to make a quantitative comparison of delay spectra obtained with MWA data and our simulations, we consider the uncertainty in the assumed spectral index of our foreground model. Our foreground models are derived from other higher frequency catalogs and sky maps. The inherent spread in spectral index increases the uncertainty while predicting fluxes at the observed frequency. Using simple error propagation, the fractional error in the delay spectrum caused by the spread in spectral index is $\sim \ln(f_\textrm{orig}/f)\,\Delta\alpha_\textrm{sp}$, where, $f_\textrm{orig}$ is the original frequency at which the catalog or map was created, $f=185$~MHz is the observed frequency, and $\Delta\alpha_\textrm{sp}$ is the spread (HWHM) in spectral index. From \citet{mau03}, we assume $\Delta\alpha_\textrm{sp} \approx 0.35$ for compact objects from NVSS and SUMSS catalogs. Although the model of \citet{deo08} yields a spectral index per direction on the sky, we assume similar uncertainties exist in spectral indices of our diffuse sky model as well, which is predominantly derived from the 408~MHz map of \citet{has82}. Thus, fractional errors in delay spectrum from compact foreground objects and diffuse emission are $\sim$70\% and $\sim$30\% respectively.

@@

Figure~\ref{fig:data-sim-ratio} shows normalized histograms of the distribution of $\log_{10}\rho$ for the {\it off--zenith} (top panel) and {\it zenith} (bottom panel) pointings respectively inside the {\it foreground wedge}. The median absolute deviation from these distributions correspond to $\sim$~90\% fractional difference between data and modeling on average in either case. This is in line with aforementioned uncertainty in foreground models and thermal noise fluctuations in measurements. Currently we are significantly limited by unavailability of all--sky foreground models at the observed frequency of 185~MHz.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Normalized histograms of logarithm of ratios of observed to simulated visibilities restricted to the {\it foreground wedge} -- i.e., within the white lines of Figure \ref{fig:sim_data} -- for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Median absolute deviation of $\log_{10}\rho$ is $\sim 0.28$ in both cases. A value of zero indicates when the ratio $\rho=1$. \label{fig:data-sim-ratio}}
% \includegraphics[width=\linewidth]{figures/v1_0/delta_array_histogram_wedge_sim_data_snr_log_ratio_0.3m_ground_custom_gaussian_FG_model_asm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
% \caption{Significance of logarithm of ratio of observed to simulated visibilities relative to expected uncertainty restricted to the {\it foreground wedge} for the {\it off--zenith} (top) and {\it zenith} (bottom) pointings. Probability densities and cumulative distributions are shown in black and gray respectively. Region inside the dotted vertical lines denotes when the ratio, $\rho$, is consistent with unity subject to expected uncertainty. Also indicated in percentages are confidence levels estimated conservatively for such a consistency. They are $\sim$~90\% and $\sim$~95\% for {\it off--zenith} and {\it zenith} pointings respectively. \label{fig:data-sim-ratio-snr}}
\end{figure}

These uncertainties are presented only to confirm the qualitative agreement already seen between Figures~\ref{fig:fhd_data} and \ref{fig:sim_data} and are not intended to serve as a comprehensive estimate of all uncertainties involved. However, since we have not accounted for numerous other uncertainties, especially in the data, such as uncertainties in the antenna power--pattern, calibration, interference, and wide field imaging effects, our estimates of uncertainty are conservative. Accounting for such uncertainties will only improve our confidence level in this comparison. Our primary objective is to explore in detail the foreground signatures embedded in the {\it foreground wedge}.

@@

Having established that the simulation matches the data to the level of expected uncertainties, we proceed to examine in further detail the key signatures seen in simulated delay spectra. A number of factors are responsible for the characteristics noted in the delay spectra obtained from data and through simulations. We address these factors below:

@@

\item {\it Sky Model}: Our model of the sky consists of bright compact objects and diffuse emission on diverse spatial scales as shown in Figures~\ref{fig:DSM} and \ref{fig:CSM}. It also consists of localized regions of strong emission such as the Galactic plane and Galactic center. The resulting sky emission is anisotropic. In fact, the patch of sky for MWA observations is chosen from regions of low foreground emission. % In our detailed study, we have divided the transiting sky into four different ``bow-tie'' shaped regions.  Each region has equal area and the regions are aligned to cardinal directions around zenith as the origin. This results in the ability to isolate features in the delay spectrum to eight different sectors of the sky.

\item {\it Baseline Orientation}: Since the spatial structure of our foreground model is not expected to be isotropic, we divide our antenna spacings by their orientation, $\theta_\textrm{b}$. We use the following bins: $-67\fdg 5\le \theta_\textrm{b} < -22\fdg 5$, $-22\fdg 5\le \theta_\textrm{b} < 22\fdg 5$, $22\fdg 5\le \theta_\textrm{b} < 67\fdg 5$, and $67\fdg 5\le \theta_\textrm{b} < 112\fdg 5$. The bin centers are oriented towards South--East, East, North--East, and North respectively. Since delays depend on $\theta_\textrm{b}$, binning by $\theta_\textrm{b}$ allows us to match delay spectrum features to different sky directions.

\item {\it Tile Pointing and Power--Pattern}: Since the MWA tiles are steered electronically, the power--pattern of the tile changes in any observing mode that tracks the object. The pointing system is capable of steering to points on a regular $\sim$7\arcdeg grid. During observations we allow the sky to drift across the nearest available pointing, shifting between grid points as necessary. In our study, we take into account the effect of changing tile power--pattern on delay spectrum (Figure~\ref{fig:power_pattern}). The LST of our data ranges from $\sim$~21 hours through $\sim$~2 hours, and for our analysis we chose {\it off--zenith} and {\it zenith} pointings shown in Figure~\ref{fig:pointings}.

\item {\it Instrumental Passband}: The instrumental passband characteristics have a direct effect on the delay transform owing to the Fourier relationship between the two. Due to excision of aliased edge channels repeated in each coarse channel of our passband, we observed repeated patterns of the {\it foreground wedge} in the delay spectra. As already mentioned earlier, we have used a deconvolution algorithm ({\it CLEAN}) to rid the delay spectra of band shape effects. Since this paper focuses on studying the effects of foreground on the delay spectra, we do not pursue the characterization of artifacts from the band shape and imperfections in {\it CLEAN} deconvolution.

@@

It may be noted that numerous features overlap at varying levels of significance depending on combinations of parameters. We assign the features to their predominant causes. Secondly, we have used noiseless cases to clearly illustrate the observed foreground signatures. With the addition of noise in the visibilities which is subject to observing time, some of the weaker features may not be as prominently visible. Since the foreground signatures are far too numerous and they reside in a vast volume of parameter space such as antenna spacing length and orientation, power--pattern, patch of sky under observation, and instrumental configuration, we highlight only some examples of the most notable features observed in the delay spectra of foreground emission.

@@

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_dsm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the diffuse foreground model without any thermal noise. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and color scale are identical to those in Figures~\ref{fig:fhd_data} and \ref{fig:sim_data}. In the {\it zenith} pointing, delay spectrum from diffuse emission has a {\it two--pronged fork}--shaped structure and is present even at wide antenna spacings. \label{fig:noiseless-dsm-delay-spectrum}}
\end{figure}

@@

The most prominent signature seen in the {\it off--zenith} pointing (top panel of Figure~\ref{fig:noiseless-dsm-delay-spectrum}) is due to the bright Galactic center situated on the western horizon co--located with one of the bright secondary lobes of the power--pattern. It appears as a bright branch in the delay spectrum near the negative delay horizon delay limit. This feature is strongest at short antenna spacings and fades with increase in antenna spacing. This Galactic center signature is absent in the {\it zenith} pointing (bottom panel) because the Galactic center has already set over the horizon.

@@

Figure~\ref{fig:noiseless-dsm-delay-spectrum} also demonstrates diffuse emission from within the primary field of view manifests in the {\it off--zenith} pointing as a branch at the top of the wedge, and at zero delay in the {\it zenith} pointing. The former is seen at positive delays because the primary lobe of the power--pattern is centered eastward of zenith, whereas in the latter it is centered at zenith. As we see from Equation \ref{eqn:obsvis}, each baseline measures a single spatial mode on the sky with an angular size scale inversely proportional to length of the baseline projected in the direction of emission. Thus, in the {\it zenith} pointing, the signature of the smooth sky model, the horizontal line at zero delay, fades away on antenna spacings $|\overline{\mathbf{x}}| \gtrsim 125$~m because the sky model is devoid of spatial structures on scales $\lesssim$~0\fdg 75. In the off-zenith pointing, we see power out to much longer baselines which is described below.

@@

A very interesting signature of diffuse emission is revealed at regions near the positive horizon delay limit in the {\it off--zenith} pointing and at both positive and negative horizon delay limits in the {\it zenith} pointing even on wide antenna spacings (125~m~$\lesssim |\overline{\mathbf{x}}|\le$~200~m). This is contrary to expectations because longer baselines are not expected to be sensitive to diffuse emission. In the wide--field observations of the MWA, long baselines appear shortened due to projection effects for off--axis emission, especially towards the horizon.  This effect is evident at all LSTs in our simulations and it naturally holds for short baselines, as well. Thus, diffuse emission from far off--axis directions manifests as an edge--heavy {\it two--pronged fork} across all baselines. It decreases in strength with increasing baseline length as expected but is present in all baseline orientations.   %% This contributes to a {\it pitchfork}--shaped feature which will be discussed later in \S\ref{sec:composite} in more detail.

@@

Figure~\ref{fig:noiseless-csm-delay-spectrum} shows the delay spectrum of compact foreground sky model (see Figure~\ref{fig:CSM})for {\it off--zenith} and {\it zenith} pointings in the top and bottom panels respectively.

\begin{figure}[htb]
\centering
\includegraphics[width=\linewidth]{figures/v1_0/delta_array_multi_baseline_CLEAN_noiseless_visibilities_0.3m_ground_custom_gaussian_FG_model_csm_all_sky_nside_64_Tsys_95.0K_185.0_MHz_30.7_MHz_bhw2.0.eps}
\caption{Simulated delay spectrum amplitudes $|V_\eta(|\overline{\mathbf{x}}|,\eta)|$ (in units of Jy~Hz) for {\it off--zenith} (top) and {\it zenith} (bottom) pointings for the compact objects foreground model without any thermal noise. White lines mark the boundaries of {\it foreground wedge} determined by the horizon delay limit and antenna spacing. The axes and logarithmic color scale are identical to those in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data} and \ref{fig:noiseless-dsm-delay-spectrum}. Emission from compact objects gives rise to centrally concentrated features and stays flat across antenna spacings in the {\it foreground wedge}.\label{fig:noiseless-csm-delay-spectrum}}
\end{figure}

@@

The amplitude response of an interferometer to a compact object is, to first order, flat with respect to baseline length. Since the main lobe of the power--pattern in the {\it off--zenith} pointing is centered eastward of zenith, the bulk of the compact foreground emission is seen in a branch with positive delay corresponding to the position of the primary lobe of the power--pattern. In the {\it zenith} pointing, compact emission from the same patch of sky is seen as a bright horizontal arm at zero delay  since the primary lobe of the power--pattern is centered at zenith.

Foreground emission in zero delay and negative delay regions in the {\it off--zenith} pointing is caused by compact objects co--located with secondary lobes of the power--pattern. On the other hand, compact objects co--located with secondary lobes of power--pattern in the {\it zenith} pointing are revealed as faint but distinct branches at positive and negative delays depending on the orientation of antenna spacing and direction of emission on the sky.

@@

@@

Figure~\ref{fig:pitchfork-baselines} shows delay spectra of three antenna pairs of different antenna spacings oriented northward during the {\it zenith} pointing; each is a different vertical slice of the two dimensional delay spectrum plots shown in Figures~\ref{fig:fhd_data}, \ref{fig:sim_data}, \ref{fig:noiseless-dsm-delay-spectrum} and \ref{fig:noiseless-csm-delay-spectrum}. The diffuse, compact, and composite components are are shown as dotted, dashed, and solid lines respectively. The horizon delay limits are shown as a pair of vertical dot--dashed lines. The gray shaded area denotes the envelope of expected uncertainty in the delay spectrum. Uncertainty in emission in the {\it foreground} wedge (between the horizon delay limits) is dominated by the $\gtrsim$~70\% error owing to uncertainties in predicting the spectral index of compact foreground objects, while thermal noise fluctuations dominate outside.
